---
title: "Assignment 2: Google Trends Data"
format: html
---

## Manual CSV Google Trends

I downloaded the CSV from Google Trends for the terms "Trump, Kamala, Election" (past 12 months).  
The CSV provides weekly intervals of search interest for each term.

```{r}
# Load manual CSV
manual_df <- read.csv("assignment02/TrumpHarrisElection.csv", skip = 2)

# Rename columns
colnames(manual_df) <- c("Week", "Trump", "Harris", "Election")

# Convert Week to Date
manual_df$Week <- as.Date(manual_df$Week)

# Replace "<1" with 0 and convert to numeric / cleaning data
manual_df$Trump    <- as.numeric(gsub("<1", "0", manual_df$Trump))
manual_df$Harris   <- as.numeric(gsub("<1", "0", manual_df$Harris))
manual_df$Election <- as.numeric(gsub("<1", "0", manual_df$Election))

# Preview
head(manual_df)

# Plotting CSV data
library(ggplot2)
ggplot(manual_df, aes(x = Week)) +
  geom_line(aes(y = Trump, color = "Trump"), linewidth = 1) +
  geom_line(aes(y = Harris, color = "Harris"), linewidth = 1) +
  geom_line(aes(y = Election, color = "Election"), linewidth = 1) +
  labs(title = "Google Trends (Manual CSV): Trump, Harris, Election",
       x = "Date", y = "Search Interest") +
  scale_color_manual(values = c("Trump"="red","Harris"="blue","Election"="orange")) +
  theme_minimal()
```

---

# gtrendsR API Google Trends

Here I collected the same data programmatically using the `gtrendsR` package.

```{r}
library(gtrendsR)
library(ggplot2)

# getting real-time data
api_res <- tryCatch({
  gtrends(
    c("Trump","Harris","Election"),
    onlyInterest = TRUE,
    geo = "US",
    time = "today 12-m"
  )
}, error = function(e) {
  message("API failed. Loading saved CSV instead.")
  return(NULL)
})

# If API call worked, extract and save
if (!is.null(api_res)) {
  the_df <- api_res$interest_over_time
  write.csv(the_df, "assignment02/TrumpHarrisElection_API.csv", row.names = FALSE)
} else {
  the_df <- read.csv("assignment02/TrumpHarrisElection_API.csv")
}

# Preview
head(the_df)

# Plot API data
# Personal note:The API data (the_df) is in "long" format because gtrendsR returns a tidy dataset
ggplot(the_df, aes(x = date, y = hits, color = keyword)) +
  geom_line(linewidth = 1) +
  labs(title = "Google Trends (API): Trump, Harris, Election",
       x = "Date", y = "Search Interest") +
  theme_minimal()
```

---

# Comparison

- **Manual CSV:**  
  - Weekly data with separate columns for each keyword.  
  - Easy to download but more to clean.
  - Not reproducible and must redownload new csv each time for updates.

- **API (gtrendsR):**  
  - More "real time" data when ran. 
  - More reproducible and customizable (timeframe, region, search property).  
  - Needs internet and may sometimes fail if Google blocks requests.  

Overall, the manual CSV is quick for a one-time analysis, while the API method is better for reproducibility and automation.

# Disclaimer
I used ChatGPT-5 to help me create an outline for this assignment and to better understand the steps involved. Since this was my first time working with data analysis in R, the outline provided useful guidance and helped me follow the process more confidently.

Prompt:

Given this assignment description and the provided CSV file, create an outline of the steps I need to take to complete the assignment. The outline should help me understand the workflow, including loading and cleaning the CSV, analyzing the data, using the gtrendsR API, and comparing the two methods.

Outline provided by ChatGPT5:

Assignment 2: Google Trends Data — Step-by-Step Outline
Set Up
Review the assignment instructions.
Make sure the CSV file (TrumpHarrisElection.csv) is saved in the correct folder (assignment02).
Open RStudio or Quarto to begin working.
Method 1: Manual CSV Data
Load the CSV into R using read.csv() (with skip to remove metadata rows).
Clean the dataset:
Rename columns for clarity.
Convert the date column into Date format.
Replace <1 values with 0 and convert to numeric.
Preview the cleaned dataset with head().
Plot search interest for Trump, Harris, and Election on the same graph using ggplot2.
Method 2: gtrendsR API Data
Install and load the gtrendsR package.
Query Google Trends for the same three terms over the last 12 months.
Extract the interest_over_time dataset.
Save the API results as a .csv and .RData file for reproducibility.
Preview the dataset with head().
Plot the trends by keyword with ggplot2.