---
title: "Education and Crime: Interactive Plots"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: false
    self_contained: true
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,   # <--- hide code in HTML
  message = FALSE,
  warning = FALSE,
  fig.width = 9,
  fig.height = 5.4
)
```

# Overview

This report explores how educational outcomes (Graduation Rate and College/Career/Military Readiness — **CCMR**) relate to **violent** and **property** crime across Dallas County districts.

- **ViolentRate** and **PropertyRate** = crime per 1,000 residents  
- **PovertyRate** = percent in poverty  

All visualizations are **interactive Plotly charts**.  
Regression models are shown using **`lm()`** and **`sjPlot::tab_model()`**.

---

# Packages

```{r packages}
pkgs <- c(
  "tidyverse", "ggplot2", "plotly",
  "viridisLite", "scales", "readr", "sjPlot"
)
to_install <- pkgs[!pkgs %in% rownames(installed.packages())]
if (length(to_install)) install.packages(to_install, repos = "https://cloud.r-project.org")

library(tidyverse)
library(ggplot2)
library(plotly)
library(viridisLite)
library(scales)
library(readr)
library(sjPlot)
```

---

# Data Load & Cleaning

```{r data}
df <- readr::read_csv("FinalData.csv", show_col_types = FALSE)

required_cols <- c(
  "Name", "GraduationRates", "CCMR.Grads",
  "ViolentRate", "PropertyRate", "PovertyRate", "TotalPopulation"
)

missing_cols <- setdiff(required_cols, names(df))
if (length(missing_cols) > 0) {
  stop(sprintf("Missing expected columns: %s", paste(missing_cols, collapse = ", ")))
}

df <- df |>
  filter(
    !is.na(GraduationRates),
    !is.na(`CCMR.Grads`),
    !is.na(ViolentRate),
    !is.na(PropertyRate)
  ) |>
  mutate(
    District = Name,
    `Poverty (%)` = PovertyRate,
    Pop = TotalPopulation,
    `Violent Crime Rate` = ViolentRate,
    `Property Crime Rate` = PropertyRate,
    `Grad Rate (%)` = GraduationRates,
    `CCMR (%)` = `CCMR.Grads`,
    hover_txt = paste0(
      "<b>", District, "</b>",
      "<br>Grad Rate: ", round(`Grad Rate (%)`, 1), "%",
      "<br>CCMR: ", round(`CCMR (%)`, 1), "%",
      "<br>Violent Crime Rate: ", round(`Violent Crime Rate`, 2), " per 1,000",
      "<br>Property Crime Rate: ", round(`Property Crime Rate`, 2), " per 1,000",
      "<br>Poverty: ", round(`Poverty (%)`, 1), "%",
      "<br>Population: ", scales::comma(Pop)
    )
  )
```

---

# Interactive Plots

## 1. Graduation vs Violent Crime Rate

```{r p_interactive}
p <- ggplot(
  df,
  aes(
    x = `Violent Crime Rate`,
    y = `Grad Rate (%)`,
    text = hover_txt,
    color = `Poverty (%)`
  )
) +
  geom_point(size = 3, alpha = 0.85) +
  geom_smooth(method = "lm", se = TRUE, alpha = 0.15) +
  scale_color_gradientn(colors = viridis(256), name = "Poverty (%)") +
  labs(
    title = "Graduation Rate vs Violent Crime Rate",
    x = "Violent crime rate (per 1,000 residents)",
    y = "Graduation Rate (%)"
  ) +
  theme_minimal(base_size = 13)

ggplotly(p, tooltip = "text")
```

---

## 2. CCMR vs Property Crime Rate

```{r q_interactive}
q <- ggplot(
  df,
  aes(
    x = `Property Crime Rate`,
    y = `CCMR (%)`,
    text = hover_txt,
    color = `Poverty (%)`
  )
) +
  geom_point(size = 3, alpha = 0.85) +
  geom_smooth(method = "lm", se = TRUE, alpha = 0.15) +
  scale_color_gradientn(colors = viridis(256), name = "Poverty (%)") +
  labs(
    title = "CCMR vs Property Crime Rate",
    x = "Property crime rate (per 1,000 residents)",
    y = "CCMR (%)"
  ) +
  theme_minimal(base_size = 13)

ggplotly(q, tooltip = "text")
```

---

## 3. Correlation Heatmap

```{r heatmap_interactive}
corr_df <- df |>
  dplyr::select(
    `Grad Rate (%)`,
    `CCMR (%)`,
    `Violent Crime Rate`,
    `Property Crime Rate`,
    `Poverty (%)`
  )

corr_mat <- round(cor(corr_df, use = "pairwise.complete.obs"), 2)

corr_long <- as.data.frame(as.table(corr_mat))
colnames(corr_long) <- c("Var1", "Var2", "Correlation")

corr_long <- corr_long |>
  mutate(
    hover_txt = paste0(
      "<b>", Var1, "</b> vs <b>", Var2, "</b>",
      "<br>Correlation: ", sprintf("%.2f", Correlation)
    )
  )

heatmap <- ggplot(
  corr_long,
  aes(
    x = Var1,
    y = Var2,
    fill = Correlation,
    text = hover_txt
  )
) +
  geom_tile(color = "white") +
  scale_fill_gradient2(
    low = "blue",
    mid = "white",
    high = "red",
    midpoint = 0,
    name = "Correlation"
  ) +
  geom_text(aes(label = sprintf("%.2f", Correlation)), size = 3) +
  coord_equal() +
  theme_minimal(base_size = 13) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(face = "bold")
  ) +
  labs(
    title = "Correlation Heatmap: Education, Crime, and Poverty",
    x = NULL,
    y = NULL
  )

ggplotly(heatmap, tooltip = "text")
```

---

## 4. Property Crime Rate vs Violent Crime Rate

```{r r_interactive}
r <- ggplot(
  df,
  aes(
    x = `Property Crime Rate`,
    y = `Violent Crime Rate`,
    text = hover_txt,
    color = `Poverty (%)`
  )
) +
  geom_point(size = 3, alpha = 0.85) +
  scale_color_gradientn(colors = viridis(256), name = "Poverty (%)") +
  labs(
    title = "Property vs Violent Crime Rates",
    x = "Property crime rate (per 1,000 residents)",
    y = "Violent crime rate (per 1,000 residents)"
  ) +
  theme_minimal(base_size = 13)

ggplotly(r, tooltip = "text")
```

---

# Regression Models

Your regression code from the prompt, unchanged and using the same CSV:

```{r regression_models}
#Install needed packages if necessary
library(tidyverse)
library(sjPlot)

dat <- read.csv("FinalData.csv")

### Regression Models
v1 <- lm(ViolentRate ~ CCMR.Grads + GraduationRates + PovertyRate + 
           MobilityRate + heterogeneity6, data = dat)
tab_model(v1)

v2 <- lm(ViolentRate ~ CCMR.College + GraduationRates + PovertyRate + 
           MobilityRate + heterogeneity6, data = dat)
tab_model(v2)

v3 <- lm(ViolentRate ~ TSI.Both + TSI.Either + DualCredit + APBI + Associates + 
           OnRamps + GraduationRates + PovertyRate + MobilityRate + heterogeneity6, 
         data = dat)
tab_model(v3)

p1 <- lm(PropertyRate ~ CCMR.Grads + GraduationRates + PovertyRate + 
           MobilityRate + heterogeneity6, data = dat)
tab_model(p1)

p2 <- lm(PropertyRate ~ CCMR.College + GraduationRates + PovertyRate + 
           MobilityRate + heterogeneity6, data = dat)
tab_model(p2)

p3 <- lm(PropertyRate ~ TSI.Both + TSI.Either + DualCredit + APBI + Associates + 
           OnRamps + GraduationRates + PovertyRate + MobilityRate + heterogeneity6, 
         data = dat)
tab_model(p3)
```

---

## Regression Visualization (Interactive, Your Style)

```{r regression_viz_interactive}
g <- ggplot(dat, aes(x = CCMR.Grads, y = ViolentRate)) + 
  geom_point(shape = 23, size = 2.25, fill = "#05f9ff", col = "#00241B") + 
  geom_smooth(method = "loess", col = "#ed1313") + 
  labs(
    x = "% of Graduates meeting at least 1 CCMR Criteria", 
    y = "Violent Crime Rates per 1000",
    title = "Violent Crime vs CCMR Graduates (Loess Fit)"
  ) + 
  theme(text = element_text(family = "serif", size = 12.5))

ggplotly(g)
```

---

# Takeaways

- **CCMR does not show a strong or consistent relationship with violent or property crime in this dataset.** The scatterplots and regression models indicate little to no association once other factors are considered.

- **Poverty has the strongest and most consistent connection to both crime and educational outcomes.** Districts with more poverty usually have more crime and lower student performance.

- **Variables like poverty and how often families move matter more for crime than CCMR or graduation rates.**
